% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}      

First and foremost, I would like to thank my supervisor, Richard E.~Turner. Both of us took risks back in summer 2013: Rich just started his faculty career and was assembling his first team, and I was very new to the machine learning world. I will never forget the struggle in the first two years when I was desperate to make my algorithms work. Hence I want to express my deepest gratitude to Rich for his detailed guidance and patience until the ``eureka'' moment in spring 2015. Also I wouldn't have done many exciting projects later on if without his encouragement of exploration and generous allowance for independent research. So now looking back to the start of this journey, I realise that I was very fortunate to have decided to join Rich's group, and also the amazing Cambridge MLG/CBL.

I am indebted to my collaborators for their supports. Jos{\'e} Miguel Hern{\'a}ndez-Lobato co-worked with me on all the experimental details in Chapter \ref{chap:factor_tying} which largely serves as the basis of the first part. I learned a lot from Qiang Liu during the collaboration in my final year, which leads to plenty of inspiration in the second part of this thesis. Collaborating with Yarin Gal has quite substantial influence on my view of Bayesian deep learning. Special thanks also go to my co-authors during my PhD study: Thang Bui, Daniel Hern{\'a}ndez-Lobato, Cuong Nguyen and Mark Rowland, all of them are knowledgable people to collaborate. 

I also want to thank a few senior researchers in particular. I had very inspiring discussions with Zoubin Ghahramani and Sebastian Nowozin during my PhD viva, and I also appreciate Zoubin's efforts to constantly bringing in great speakers, and the precious schedule of 1-to-1 meeting with them. Martin Szummer and Ulrich Paquet, when they were both in Cambridge, provided advice when I was grinding message passing algorithms in the first two years. Stephan Mandt hosted my internship at Disney Research in Jun-Sept 2017, where I had a great summer there for both exciting research projects and enjoyable activities, including my first heavy metal concert.

%
I would like to thank all I have met in Cambridge for comfortable leisure time. I also thank the FFTF fellowship award from Schlumberger Foundation (2014, 2015, 2016). 

I can never thank my family enough. My father kept reminding me to read books. My mother concerned a lot about my body health and leisure life. My sister Yinghong shared with me the joy and sorrow of tuning hyper-parameters of deep neural networks.

\end{acknowledgements}
