% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.

\begin{publication}
Some of the work in this thesis has been published in the following venues.

Section \ref{sec:chap2_ep_energy} adapts from:
\begin{itemize}
\item \textbf{Yingzhen Li} and Richard E.~Turner. \\
\emph{A Unifying Approximate Inference Framework from Variational Free Energy Relaxation}. \\
Advances in Approximate Inference workshop, NIPS, 2016.
\end{itemize}

Chapter \ref{chap:factor_tying} expands on:
\begin{itemize}
\item \textbf{Yingzhen Li}, Jos{\'e} Miguel Hern{\'a}ndez-Lobato and Richard E.~Turner. \\
\emph{Stochastic Expectation Propagation}. \\
Neural Information Processing Systems (NIPS), 2015.

%\item Jos{\'e} Miguel Hern{\'a}ndez-Lobato*, \textbf{Yingzhen Li}*, Mark Rowland, Daniel Hern{\'a}ndez-Lobato, Thang Bui and Richard E.~Turner. \\
%\emph{Black-box $\alpha$-Divergence Minimization}. \\
%International Conference on Machine Learning (ICML), 2016. (joint first author)

\end{itemize}

Chapter \ref{chap:vrbound} expands on:

\begin{itemize}
\item \textbf{Yingzhen Li} and Richard E.~Turner. \\
\emph{R{\'e}nyi Divergence Variational Inference}. \\
Neural Information Processing Systems (NIPS), 2016.

\end{itemize}

Chapter \ref{chap:wild} expands on:

\begin{itemize}
\item \textbf{Yingzhen Li} and Qiang Liu. \\
\emph{Wild Variational Approximations}. \\
Advances in Approximate Inference workshop, NIPS, 2016.
\end{itemize}

%Chapter \ref{chap:dropout_alpha} expands on:
%\begin{itemize}
%\item \textbf{Yingzhen Li} and Yarin Gal. \\
%\emph{Dropout Inference for Bayesian Neural Networks with Alpha-divergences}. \\
%International Conference on Machine Learning (ICML), 2017.
%\end{itemize}

Chapter \ref{chap:grad_approx} expands on:
\begin{itemize}
\item \textbf{Yingzhen Li} and Richard E.~Turner. \\
\emph{Gradient Estimators for Implicit Models}. \\
International Conference on Learning Representations (ICLR), 2018.
\end{itemize}

%Chapter \ref{chap:amcmc} expands on:
%\begin{itemize}
%\item \textbf{Yingzhen Li}, Richard E.~Turner and Qiang Liu. \\
%\emph{Approximate Inference with Amortised MCMC}. \\
%In submission to NIPS 2017.
%\end{itemize}

\end{publication}
